Real-space computation of $E$/$B$-mode maps I: Formalism, Compact Kernels, and Polarized Filaments

### 2018-07-31 articles/formalism/sections/kernel_paintbrush.tex
Added clarifying sentence on direction in plots.

The $B$ part of the Green's function, traced along the filament, cancels itself except at the filament ends.  This results in a non-zero $B$ pattern for the filament.  
 For a North--South filament, the $B$-mode pattern is positive on the North-East and South-West, and negative in the North-West and South-East as seen in \fig{fig:polfilaments}.  The size of this $B$-mode pattern is set by the dimensions of the filament, chiefly the width.  In contrast to the kernel radial fuctions, which are more compact at higher band limit, the filament $B$-mode  pattern does not depend much on the band limit, provided it is sufficient to resolve the filament structure.  A filament with a more gradual edge has a $B$-pattern that is more spread out, but follows the same general structure.

@@

Changing the polarization direction within the filament changes the $E/B$ patterns.  A $90^\circ$ rotation of the polarization with respect to the filament changes the sign of both $E$ and $B$.  In a polarization pattern aligned at $45^\circ$ to the filament, the $E$ pattern will swap with the $B$ pattern.  The work in \cite{Zaldarriaga2001a} correctly argued that linear filaments (infinite and without ends) can only produce $E$-power, but we find that for realistic filaments, the ends produce $B$-mode power with a clear signature.  Careful study of the $E/B$ mode power in filaments can provide insights into the orientation of the magnetic field with respect to the axis of the filament, this information could potentially shed light on the internal dynamics of filaments.
% 
\begin{figure}[t]
\includegraphics[width=0.5\columnwidth]{line.pdf}
\includegraphics[width=0.5\columnwidth]{spiral.pdf}
\includegraphics[width=1.0\columnwidth]{colorbar.pdf}
\caption{ The polarization signals of toy filament structures. In a filament organized perfectly along a magnetic field line, the polarization will be perpendicular to the filament direction.  The $E/B$ modes of filaments are in some ways easier to think about than the Stokes parameters. Left panels: in a straight filament, the $E$-mode is positive along the filament and at the ends, but negative along the sides.  The $B$-modes are only non-zero at the ends.  Right panels: in a curved filament, the $E$-mode is again positive along the filament.  Outside the filament, the $E$-mode is more negative on the interior of the curve than the exterior.  The $B$-modes are again non-zero at the ends, akin to the straight filament case, but also along the filament due to the changing radius of curvature. In all images, the longitude angle increases to the left (which is East in sky convention).  All plot are on a common, arbitrary color scale.}
\label{fig:polfilaments}
\end{figure}
%
\indent The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern. Here it is important to note that this cancellation will not happen for an ellipse or any loop of non-constant radius of curvature.
The same general description holds for a spiral-shaped filament, which can can be viewed as the distortion of the straight filament (right panel of  \fig{fig:polfilaments}).  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside. While the $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations, note that it is non-vanishing in the intermediate regions along the filament due to the varying radius of curvature of the spiral.  A spiral turning with the other handedness will have a $B$-pattern with the opposite sign.

>>>>

The $B$ part of the Green's function, traced along the filament, cancels itself except at the filament ends.  This results in a non-zero $B$ pattern for the filament.  
 For a North--South filament, the $B$-mode pattern is positive on the northeast and southwest, and negative in the northwest and southeast. (\fig{fig:polfilaments} shows East to the left in sky map convention.)  The size of this $B$-mode pattern is set by the dimensions of the filament, chiefly the width.  In contrast to the kernel radial fuctions, which are more compact at higher band limit, the filament $B$-mode  pattern does not depend much on the band limit, provided it is sufficient to resolve the filament structure.  A filament with a more gradual edge has a $B$-pattern that is more spread out, but follows the same general structure.  The work in \cite{Zaldarriaga2001a} correctly argued that linear filaments that are infinite and without ends can only produce $E$-power, but we find that for realistic filaments, the ends produce $B$-mode power with a clear signature.

@@

 Changing the polarization direction within the filament changes the $E/B$ patterns.  A $90^\circ$ rotation of the polarization with respect to the filament changes the sign of both $E$ and $B$.  In a polarization pattern aligned at $45^\circ$ to the filament, the $E$ pattern will swap with the $B$ pattern.  Careful study of the $E/B$ mode power in filaments can provide insights into the orientation of the magnetic field with respect to the axis of the filament, this information could potentially shed light on the internal dynamics of filaments.
% 
\begin{figure}[t]
\includegraphics[width=0.5\columnwidth]{line.pdf}
\includegraphics[width=0.5\columnwidth]{spiral.pdf}
\includegraphics[width=1.0\columnwidth]{colorbar.pdf}
\caption{ The polarization signals of toy filament structures. In a filament organized perfectly along a magnetic field line, the polarization will be perpendicular to the filament direction.  The $E/B$ modes of filaments are in some ways easier to think about than the Stokes parameters. Left panels: in a straight filament, the $E$-mode is positive along the filament and at the ends, but negative along the sides.  The $B$-modes are only non-zero at the ends.  Right panels: in a curved filament, the $E$-mode is again positive along the filament.  Outside the filament, the $E$-mode is more negative on the interior of the curve than the exterior.  The $B$-modes are again non-zero at the ends, akin to the straight filament case, but also along the filament due to the changing radius of curvature. In all images, the longitude angle increases to the left (which is East in sky convention).  All plot are on a common, arbitrary color scale.}
\label{fig:polfilaments}
\end{figure}

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern. Here it is important to note that this cancellation will not happen for an ellipse or any loop of non-constant radius of curvature.
The same general description holds for a spiral-shaped filament, which can can be viewed as the distortion of the straight filament (right panel of  \fig{fig:polfilaments}).  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside. While the $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations, note that it is non-vanishing in the intermediate regions along the filament due to the varying radius of curvature of the spiral.  A spiral turning with the other handedness will have a $B$-pattern with the opposite sign.
